SAM simétrico y asimétrico

### 2020-06-30 analysis/paper/paper.Rmd
Efficiency improvements

```{r read-hgt, cache = FALSE}
hgt <- ReadNetCDF(ERA5(), 
                  subset = list(time = c("1979-01-01", "2018-12-31"),
                                latitude = c(-90, 10)),
                  vars = c(hgt = "z")) %>% 
  normalise_coords() %>% 
  .[, hgt := hgt/9.8] %>% 
  .[, dataset := "ERA5"] %>%  
  .[, hgt_a := hgt - mean(hgt), by = .(lon, lat, lev, month(time), dataset)] %>% 
  .[, hgt_m := mean(hgt_a), by = .(lat, lev, time)] %>% 
  .[, hgt_z := hgt_a - hgt_m]
```

```{r compute-SAM}
SAM <- hgt[lat <= -20] %>% 
  .[, hgt := hgt_a*sqrt(cos(lat*pi/180))] %>% 
  .[, .(eof = list(eof_flip(EOF(hgt ~ time | lon + lat, n = 1, data = .SD)))), 
    by = .(dataset, lev)]
```

@@

```{r method, fig.cap = "Spatial patterns of the first EOF of 700 hPa geopotential height. Full field (left), zonally asymmetric component (middle) and zonally symmetric component (right). Arbitrary units."}
sam_sep <- SAM[, eof[[1]]$right, by = .(dataset, lev)] %>% 
  setnames("hgt", "full") %>% 
  .[, c("sym", "asym") := list(mean(full), Anomaly(full)), by = .(lat, lev, dataset)]

@@

sam_sep %>% 
  .[dataset == "ERA5"] %>% 
  .[lev == 700] %>%
  rm_singleton() %>% 
  melt(id.vars = c("lon", "lat")) %>%

@@

```{r compute-asymsam}
indexes <-  hgt[lat <= -20, eof_asym(hgt_a, lon, lat, time, n = 1), by = lev]
indexes_wide <- dcast(indexes, + lev + time ~ term, value.var = "estimate_norm")
```

>>>>

```{r read-hgt}
hgt <- ReadNetCDF(ERA5(), 
                  subset = list(time = c("1979-01-01", "2018-12-31"),
                                latitude = c(-90, 10),
                                level = as.list(main_levs)),
                  vars = c(hgt = "z")) %>% 
  normalise_coords() %>% 
  .[, hgt := hgt/9.8] %>% 
  .[, hgt_a := hgt - mean(hgt), by = .(lon, lat, lev, month(time))]
```

```{r compute-SAM}
SAM <- hgt[lat <= -20] %>% 
  .[, hgt := hgt_a*sqrt(cos(lat*pi/180))] %>% 
  .[, .(eof = list(eof_flip(EOF(hgt ~ time | lon + lat, n = 1, data = .SD)))), 
    by = .(lev)]
```

@@

```{r method, fig.cap = "Spatial patterns of the first EOF of 700 hPa geopotential height. Full field (left), zonally asymmetric component (middle) and zonally symmetric component (right). Arbitrary units."}
sam_sep <- SAM[, eof[[1]]$right, by = .(lev)] %>% 
  setnames("hgt", "full") %>% 
  .[, c("sym", "asym") := list(mean(full), Anomaly(full)), by = .(lat, lev)]

@@

sam_sep %>% 
  .[lev == 700] %>%
  rm_singleton() %>% 
  melt(id.vars = c("lon", "lat")) %>%

@@

```{r compute-asymsam}
indexes_file <- data_path("derived", "indexes.Rds")
if (file.exists(indexes_file)) {
  indexes <- readRDS(indexes_file)
} else {
  stop("Run 02-compute-eofs.R")
}

indexes_wide <- dcast(indexes, + lev + time ~ term, value.var = "estimate_norm")
```
